---
title: "Regex Commands - Titanic Dataset"
output: html_notebook
---
Understanding the Basic Regex Commands.

```{r data_load}
titanic_df <- read.csv("titanic_data.csv")
```

Structure and Dimension of the Data

```{r str_dim}
str(titanic_df)
```

Basic Overall Plots - Age, Survived, Fare [Getting Insigts]

```{r plot_a}
hist(titanic_df$Age)
```

```{r plot_b}
library(ggplot2)

ggplot(aes(x = Age), data = titanic_df ) +
  geom_histogram(aes(fill=Sex)) +
  facet_grid(.~Survived) + 
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70 ,80 ,90 , 100)) +
  ggtitle("Overall Survivability") +
  theme(plot.title = element_text(hjust = 0.5))+
  xlab('Age')
```

```{r plot_c}

ggplot(aes(x = Fare), data = titanic_df ) +
  geom_histogram(aes(fill=Sex)) +
  facet_grid(.~Pclass) + 
  ggtitle("Fare Distribution across Class") +
  theme(plot.title = element_text(hjust = 0.5))+
  xlab('Fare')
```

Using Regex Commands in Name Column:

Have a look at the Name column of titanic, that contains the names of the passengers. The name of the first passenger in the dataset is Braund, Mr. Owen Harris. All these names have a common format. First, we have the family name, next there's the title, followed by the first and middle names.

Because this layout is consistent throughout the Name column, we're able to infer the gender of the passenger: men have the title Mr., women have the title Mrs or Miss.

```{r regex}

pass_names <- titanic_df$Name

is_man <- grepl(pattern = ", Mr\\.", pass_names)

print(paste("The Pattern Sum is:" ,sum(is_man)))

print(paste("The Total Sum is:",sum(titanic_df$Sex == "male")))

```

The sum of is_man doesn't match with the Sex column in titanic, so we've must have missed some titles that also refer to men!

It appeared that the title Mr. may not cover all men on board. Instead of manually going through all titles that appear in the Name column of titanic, we can write a clever gsub() command that extracts the title part.

The pattern we'll need is the following:

**"^.*, (.*?)\\..*$"**

With ^ and $ we signify the start and end of the string. Next, we have two .* parts in there: wildcards for the last name and first names. With , (.*?)\\. we use a similar pattern as before, but the parentheses allow us to re-use whatever is matched inside the parentheses in our replacement.

```{r regex_a}
titles <- gsub(pattern = "^.*, (.*?)\\..*$", "\\1", pass_names)

unique(titles)

```

```{r regex_b}

pass_names[1:5]

```

```{r regex_c}
convert_name <- function(name) {
  # women: take name from inside parentheses
  if (grepl("\\(.*?\\)", name)) {
    gsub("^.*?\\((.*?)\\)$", "\\1", name)
  # men: take name before comma and after title
  } else {
    gsub("^(.*?),\\s[a-zA-Z\\.]*?\\s(.*?)$", "\\2 \\1", name)
  }
}

# Call convert_name on name
clean_pass_names <- vapply(pass_names, FUN = convert_name,
                           FUN.VALUE = character(1), USE.NAMES = FALSE)

# Print out clean_pass_names
titanic_df$Modified_Names <- clean_pass_names

```

